---
title: "UK General Election 2017"
author: "Toby Staines"
date: "13 December 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prepare R Environment

Load the required libraries

```{r, echo=False}
# For spatial data handling
library(rgdal)
library(spdep)
library(rgeos)
# For charting
library(tmap)
library(cartogram)
library(ggplot2)
library(gridExtra)
library(GGally)
# For data loding and munging
library(readr)
library(dplyr)
library(tidyr)
library(car)
# For spatial stats
library(GWmodel)
library(spdep)
# For cluster analysis
library(cluster)
# For political mapping
library(leaflet)
library(sf)
library(htmlwidgets)
library(parlitools)
```

## Load and Prepare Data
Load the raw data (2015 results and swing have already been incorporated in Excel):
```{r, echo=FALSE}

election17 <- read_csv('HoC-GE2017-constituency-results.csv')
census_data_ts <- read_csv('2011_census_oa_working.csv')


```

Pre-process data:

```{r, echo=TRUE, eval=FALSE}

# Group census data to constituency level and create combined variables  
census_data_ts <- census_data_ts %>%
  group_by(PCON16CD)%>%
  summarise(
    total_pop = sum(Total_Population),
    younger_adults = sum(Age_20_to_24, Age_25_to_29, Age_30_to_44) / sum(Total_Population),
    young_adults = sum(Age_20_to_24) / sum(Total_Population),
    white = sum(White_British_and_Irish) / sum(Total_Population),
    christian = sum(Christian) / sum(Total_Population),
    english_speaking = sum(Main_language_is_English_or_Main_language_not_English__Can_speak_English_very_well)
    / sum(Total_Population),
    single_ethnicity_household = sum(All_household_members_have_the_same_ethnic_group) 
    / sum(Total_Households),
    own_home = sum(Owned_and_Shared_Ownership) / sum(Total_Households),
    social_housing = sum(Social_rented) / sum(Total_Households),
    not_good_health = sum(Fair_health, Bad_health, Very_bad_health) / sum(Total_Population),
    degree_educated = sum(Highest_level_of_qualification_Level_4_qualifications_and_above) / 
      sum(Highest_level_of_qualification_Level_4_qualifications_and_above,
          Highest_level_of_qualification_Level_3_qualifications,
          Highest_level_of_qualification_Level_1_Level_2_or_Apprenticeship,
          No_qualifications),
    no_car = sum(No_cars_or_vans_in_household) / sum(Total_Households),
    private_transport_to_work = sum(Private_Transport) / sum(Total_Employment_16_to_74),
    professionals = sum(Managers_directors_and_senior_officials, Professional_occupations) /
      sum(Total_Employment_16_to_74)
  )

# Add election result data
census_data_ts <- merge(census_data_ts, election17[, c("ons_id", "conLabSwing", "labPct15", "conPct15")], by.x="PCON16CD", by.y="ons_id")

# Create hexegon map data frame of UK constituencies
west_hex_map <- parlitools::west_hex_map #Base map
west_hex_map$region_name <- gsub(" Euro Region","",west_hex_map$region_name)
west_hex_map$region_name <- gsub("North East","NE, Yorkshire and Humber",west_hex_map$region_name)
west_hex_map$region_name <- gsub("Yorkshire and the Humber","NE, Yorkshire and Humber",west_hex_map$region_name)
 
swing_map <- inner_join(west_hex_map, census_data_ts, by=c("gss_code" = "PCON16CD"))
census_data_ts <- inner_join(census_data_ts, swing_map[,c("gss_code","region_name")], by=c("PCON16CD"="gss_code"))
# Remove redundant source data
rm(election17)

```

## Visualise Electoral Success Metrics
Leaflet plots of swing and share of vote:
```{r, echo=TRUE, eval=FALSE}
# Plot map of swing from Conservative to Labour at 2017 election
pal1 = colorNumeric("RdBu", swing_map$conLabSwing, reverse=TRUE, na.color="white")

label_no1 <- paste0(
  "<strong>", swing_map$constituency_name, "</strong>", "</br>",
  "Swing: ", swing_map$conLabSwing
) %>% lapply(htmltools::HTML)

leaflet(options=leafletOptions(
  dragging = FALSE, zoomControl = FALSE, tap = FALSE,
  attributionControl = FALSE),
  swing_map) %>%
  addPolygons(
    color = "grey",
    weight=0.75,
    opacity = 0.5,
    fillOpacity = 1,
    fillColor = ~pal1(conLabSwing),
    label = label_no1) %>%
  addLegend("topright", pal = pal1, values = ~conLabSwing,
    title = "2017 Swing to Labour",
    opacity = 1)  %>% 
  htmlwidgets::onRender(
    "function(x, y) {
        var myMap = this;
        myMap._container.style['background'] = '#fff';
    }")%>% 
  mapOptions(zoomToLimits = "first")

#Plot Labout share of vote at 2015 election
pal2 = colorNumeric("Reds", swing_map$labPct15, na.color="white")

label_no2 <- paste0(
  "<strong>", swing_map$constituency_name, "</strong>", "</br>",
  "2015 Labour vote: ", swing_map$labPct15
) %>% lapply(htmltools::HTML)

leaflet(options=leafletOptions(
  dragging = FALSE, zoomControl = FALSE, tap = FALSE,
  attributionControl = FALSE),
  swing_map) %>%
  addPolygons(
    color = "grey",
    weight=0.75,
    opacity = 0.5,
    fillOpacity = 1,
    fillColor = ~pal2(labPct15),
    label = label_no2) %>%
  addLegend("topright", pal = pal2, values = ~labPct15,
    title = "2015 Labour Vote",
    opacity = 1)  %>% 
  htmlwidgets::onRender(
    "function(x, y) {
        var myMap = this;
        myMap._container.style['background'] = '#fff';
    }")%>% 
  mapOptions(zoomToLimits = "first")

```

## Investigate Variable Relationships
Plot demographics against swing:
```{r, echo=TRUE, eval=FALSE}
swing_map %>%

  gather(c(young_adults, younger_adults, degree_educated, professionals, white, christian, english_speaking,  single_ethnicity_household, social_housing, no_car, not_good_health), key = "expl_var", value="const_prop") %>%

  ggplot(aes(x=const_prop, y=conLabSwing))+
  labs(x="Proportion of Constituency",y="Swing - Conservative to Labour") +
  geom_point(aes(colour=region_name),pch=19) +
  stat_smooth(method=lm, se=FALSE, size=1, colour="black")+
  scale_color_brewer(palette="Paired")+
  facet_wrap(~expl_var, scales="free")+
  theme_bw()+
  ggtitle("Demographic Variables vs 2017 Swing")
```

PLot demographics against vote share:
```{r, echo=TRUE, eval=FALSE}
swing_map %>%

  gather(c(young_adults, younger_adults, degree_educated, professionals, white, christian, english_speaking,  single_ethnicity_household, social_housing, no_car, not_good_health), key = "expl_var", value="const_prop") %>%

  ggplot(aes(x=const_prop, y=labPct15))+ 
  labs(x="Proportion of Constituency",y="2015 Labour Vote Share") +
  geom_point(aes(colour=region_name),pch=19) +
  stat_smooth(method=lm, se=FALSE, size=1, colour="black")+
  scale_color_brewer(palette="Paired")+
  facet_wrap(~expl_var, scales="free")+
  theme_bw()+
  ggtitle("Demographic Variables vs 2015 Labour Vote Share")
```


```{r, echo=TRUE, eval=FALSE}

region_const_count <- census_data_ts %>%
  group_by("region_name") %>%
  summarise()


```

Correlation between baseline and swing:

```{r, echo=TRUE, eval=FALSE}
swing_map %>%
  ggplot(aes(x=labPct15, y=conLabSwing))+ 
  geom_point(aes(colour=region_name),pch=19) +
  labs(x="2017 Swing to Labour",y="2015 Labour Vote Share")+
  stat_smooth(method=lm, se=FALSE, size=1, colour="#525252")+
  scale_color_brewer(palette="Paired")+
  theme_bw()+
  ggtitle(paste("Correlation:",round(cor.test(swing_map$conLabSwing,swing_map$labPct15)$estimate,2)))

```

Calculate correlation between variables:
```{r, echo=TRUE, eval=FALSE}
# Generate correlation coefficient matrix
census_data_ts %>% 
  select(conLabSwing, labPct15, young_adults, younger_adults, degree_educated, professionals, white, christian, english_speaking,  single_ethnicity_household, social_housing, no_car, not_good_health) %>%
  ggcorr(label=TRUE,nbreaks=5,  geom = "text", hjust = 1, size = 3, palette = "RdYlGn", label_alpha=0.1, layout.exp = 2)

# Look again at Scotland using a call to filter().
census_data_ts %>% 
  filter(region_name == "Scotland") %>%
  select(conLabSwing, labPct15, young_adults, younger_adults, degree_educated, professionals, white, christian, english_speaking,  single_ethnicity_household, social_housing, no_car, not_good_health) %>%
  ggcorr(label=TRUE,nbreaks=5,  geom = "text", hjust = 1, size = 3, palette = "RdYlGn", label_alpha=0.1, layout.exp = 2)

# Calculate VIF scores
# For 2015 result
lab15_vif <- vif(lm(conLabSwing ~ younger_adults + white + degree_educated, data=census_data_ts))

View(lab15_vif)

# For 2017 swing
swing_vif <- vif(lm(conLabSwing ~ young_adults + younger_adults + white + english_speaking + not_good_health + degree_educated + no_car + professionals + social_housing, data=census_data_ts))

#View(swing_vif)
```

## Regression Models
Prepare regression models and plot residuals:
```{r, echo=TRUE, eval=FALSE}



m_15_multi <- lm(labPct15 ~ younger_adults + white + english_speaking + not_good_health + social_housing , data=swing_map)
m_15_nocar <- lm(labPct15 ~ no_car, data=swing_map)
m_swing_multi <- lm(conLabSwing ~ younger_adults + degree_educated + white, data=swing_map)
m_swing_degree <- lm(conLabSwing ~ degree_educated, data=swing_map)

swing_map$resids_15_multi <- resid(m_15_multi)
swing_map$resids_15_nocar <- resid(m_15_nocar)
swing_map$resids_swing_degree <- resid(m_swing_degree)
swing_map$resids_swing_multi <- resid(m_swing_multi)


norm_resids <- function(data) {
  if (abs(min(data)) > max(data)) {
       max_range_value = abs(min(data))
       min_range_value = min(data)
  } else {
    max_range_value = max(data)
    min_range_value = -max(data)
  }
  (2 * data) / (max_range_value - min_range_value)
  }

swing_map$resids_15_nocar <- norm_resids(swing_map$resids_15_nocar)
swing_map$resids_swing_degree <- norm_resids(swing_map$resids_swing_degree)
swing_map$resids_15_multi <- norm_resids(swing_map$resids_15_multi)
swing_map$resids_swing_multi <- norm_resids(swing_map$resids_swing_multi)


```
```{r, echo=TRUE, eval=FALSE}

pal1 = colorNumeric("RdBu", domain=c(-1.0,1.0), reverse=TRUE)

label_no1 <- paste0(
  "<strong>", swing_map$constituency_name, "</strong>", "</br>",
  "Residual: ", swing_map$resids_15_nocar
) %>% lapply(htmltools::HTML)

leaflet(options=leafletOptions(
  dragging = FALSE, zoomControl = FALSE, tap = FALSE,
  attributionControl = FALSE),
  swing_map) %>%
  addPolygons(
    color = "grey",
    weight=0.75,
    opacity = 0.5,
    fillOpacity = 1,
    fillColor = ~pal1(resids_15_nocar),
    label = label_no1) %>%
  addLegend("topright", pal = pal1, values = c(-1.0,1.0),
    title = "<p>Normalised Residuals from</p><p>Univariate Global Model</p>",
    opacity = 1)  %>% 
  htmlwidgets::onRender(
    "function(x, y) {
        var myMap = this;
        myMap._container.style['background'] = '#fff';
    }")%>% 
  mapOptions(zoomToLimits = "first")



label_no2 <- paste0(
  "<strong>", swing_map$constituency_name, "</strong>", "</br>",
  "Residual: ", swing_map$resids_swing_degree
) %>% lapply(htmltools::HTML)

leaflet(options=leafletOptions(
  dragging = FALSE, zoomControl = FALSE, tap = FALSE,
  #minZoom = 6, maxZoom = 6, maxBounds = list(list(2.5,-7.75),list(58.25,50.0)),
  attributionControl = FALSE),
  swing_map) %>%
  addPolygons(
    color = "grey",
    weight=0.75,
    opacity = 0.5,
    fillOpacity = 1,
    fillColor = ~pal1(resids_swing_degree),
    label = label_no2) %>%
  addLegend("topright", pal = pal1, values = c(-1.0,1.0),
    title = "<p>Normalised Residuals from</p><p>Univariate Global Model</p>",
    opacity = 1)  %>% 
  htmlwidgets::onRender(
    "function(x, y) {
        var myMap = this;
        myMap._container.style['background'] = '#fff';
    }")%>% 
  mapOptions(zoomToLimits = "first")

#summary(m_15_multi)
#vif(m_15_multi)

#summary(m_swing_multi)
#vif(m_swing_multi)

```
Repeat reression models using only England and Wales data:
```{r, echo=TRUE, eval=FALSE}
ew_swing_map <- swing_map %>%
  filter(region_name != "Scotland")

m_ew_15_nocar <- lm(labPct15 ~ no_car, data=ew_swing_map)
ew_swing_map$resids_ew_15_nocar <- resid(m_ew_15_nocar)
#ew_swing_map$resids_ew_15_nocar <- norm_resids(ew_swing_map$resids_ew_15_nocar)

m_ew_swing_degree <- lm(conLabSwing ~ degree_educated, data=ew_swing_map)
ew_swing_map$resids_ew_swing_degree <- resid(m_ew_swing_degree)
#ew_swing_map$resids_ew_swing_degree <- norm_resids(ew_swing_map$resids_ew_swing_degree)

m_ew_15_multi <- lm(labPct15 ~ younger_adults + white + english_speaking + not_good_health + social_housing, data=ew_swing_map)
ew_swing_map$resids_ew_15_multi <- resid(m_ew_15_multi)
#ew_swing_map$resids_ew_15_multi <- norm_resids(ew_swing_map$resids_ew_15_multi)

m_ew_swing_multi <- lm(conLabSwing ~ younger_adults + degree_educated + white, data=ew_swing_map)
ew_swing_map$resids_ew_swing_multi <- resid(m_ew_swing_multi)
#ew_swing_map$resids_ew_swing_multi <- norm_resids(ew_swing_map$resids_ew_swing_multi)

label_no1 <- paste0(
  "<strong>", ew_swing_map$constituency_name, "</strong>", "</br>",
  "Residual: ", ew_swing_map$resids_ew_swing_degree
) %>% lapply(htmltools::HTML)

leaflet(options=leafletOptions(
  dragging = FALSE, zoomControl = FALSE, tap = FALSE,
  attributionControl = FALSE),
  ew_swing_map) %>%
  addPolygons(
    color = "grey",
    weight=0.75,
    opacity = 0.5,
    fillOpacity = 1,
    fillColor = ~pal1(resids_ew_swing_degree),
    label = label_no1) %>%
  addLegend("topright", pal = pal1, values = c(-0.45,0.45),
    title = "<p>Normalised Residuals from</p><p>Univariate Global Model</p>",
    opacity = 1)  %>% 
  htmlwidgets::onRender(
    "function(x, y) {
        var myMap = this;
        myMap._container.style['background'] = '#fff';
    }")%>% 
  mapOptions(zoomToLimits = "first")

label_no2 <- paste0(
  "<strong>", ew_swing_map$constituency_name, "</strong>", "</br>",
  "Residual: ", ew_swing_map$resids_ew_swing_degree
) %>% lapply(htmltools::HTML)

leaflet(options=leafletOptions(
  dragging = FALSE, zoomControl = FALSE, tap = FALSE,
  attributionControl = FALSE),
  ew_swing_map) %>%
  addPolygons(
    color = "grey",
    weight=0.75,
    opacity = 0.5,
    fillOpacity = 1,
    fillColor = ~pal1(resids_ew_swing_degree),
    label = label_no2) %>%
  addLegend("topright", pal = pal1, values = c(-0.45,0.45),
    title = "<p>Normalised Residuals from</p><p>Univariate Global Model</p>",
    opacity = 1)  %>% 
  htmlwidgets::onRender(
    "function(x, y) {
        var myMap = this;
        myMap._container.style['background'] = '#fff';
    }")%>% 
  mapOptions(zoomToLimits = "first")

summary(m_ew_15_multi)
summary(m_ew_swing_multi)
```

Model each variable and get adjusted R^2:

```{r, echo=TRUE, eval=FALSE}

m_15_young <- lm(labPct15 ~ young_adults, data=swing_map)
m_15_younger <- lm(labPct15 ~ younger_adults, data=swing_map)
m_15_degree <- lm(labPct15 ~ degree_educated, data=swing_map)
m_15_prof <- lm(labPct15 ~ professionals, data=swing_map)
m_15_white <- lm(labPct15 ~ white, data=swing_map)
m_15_christian <- lm(labPct15 ~ christian, data=swing_map)
m_15_englishSp <- lm(labPct15 ~ english_speaking, data=swing_map)
m_15_singleEth <- lm(labPct15 ~ single_ethnicity_household, data=swing_map)
m_15_social <- lm(labPct15 ~ social_housing, data=swing_map)
m_15_nocar <- lm(labPct15 ~ no_car, data=swing_map)
m_15_ngh <- lm(labPct15 ~ not_good_health, data=swing_map)

m_swing_young <- lm(conLabSwing ~ young_adults, data=swing_map)
m_swing_younger <- lm(conLabSwing ~ younger_adults, data=swing_map)
m_swing_degree <- lm(conLabSwing ~ degree_educated, data=swing_map)
m_swing_prof <- lm(conLabSwing ~ professionals, data=swing_map)
m_swing_white <- lm(conLabSwing ~ white, data=swing_map)
m_swing_christian <- lm(conLabSwing ~ christian, data=swing_map)
m_swing_englishSp <- lm(conLabSwing ~ english_speaking, data=swing_map)
m_swing_singleEth <- lm(conLabSwing ~ single_ethnicity_household, data=swing_map)
m_swing_social <- lm(conLabSwing ~ social_housing, data=swing_map)
m_swing_nocar <- lm(conLabSwing ~ no_car, data=swing_map)
m_swing_ngh <- lm(conLabSwing ~ not_good_health, data=swing_map)

summary(m_15_young)$adj.r.squared
summary(m_15_younger)$adj.r.squared
summary(m_15_degree)$adj.r.squared
summary(m_15_prof)$adj.r.squared
summary(m_15_white)$adj.r.squared
summary(m_15_christian)$adj.r.squared
summary(m_15_englishSp)$adj.r.squared
summary(m_15_singleEth)$adj.r.squared
summary(m_15_social)$adj.r.squared
summary(m_15_nocar)$adj.r.squared
summary(m_15_ngh)$adj.r.squared

summary(m_swing_young)$adj.r.squared
summary(m_swing_younger)$adj.r.squared
summary(m_swing_degree)$adj.r.squared
summary(m_swing_prof)$adj.r.squared
summary(m_swing_white)$adj.r.squared
summary(m_swing_christian)$adj.r.squared
summary(m_swing_englishSp)$adj.r.squared
summary(m_swing_singleEth)$adj.r.squared
summary(m_swing_social)$adj.r.squared
summary(m_swing_nocar)$adj.r.squared
summary(m_swing_ngh)$adj.r.squared
```

## Globally Weighted Summary Statistics
Prepare GWSS data:
```{r, echo=TRUE, eval=FALSE}

pc_boundaries <- readOGR(dsn = "GB_Hex_Cartogram_Const", layer = "GB_Hex_Cartogram_Const")
pc_boundaries@data <- left_join(pc_boundaries@data, census_data_ts, by=c("CODE"="PCON16CD"))

pc_boundaries <- pc_boundaries[!is.na(pc_boundaries@data$labPct15),]

# Note that we use the 'optimal' bandwidth derived from the bw.gwr() function.
gw_ss_15 <- gwss(pc_boundaries, vars  =  c("labPct15", "conLabSwing", "younger_adults", "degree_educated",
                                           "professionals", "white", "christian", "english_speaking",
                                           "single_ethnicity_household", "social_housing", "no_car", "not_good_health"), kernel =
                   "bisquare", adaptive = TRUE, bw = 75, quantile = TRUE)

gw_ss_15_cols <- as.data.frame(colnames(gw_ss_15$SDF@data))
```
## Sepearate GW plots

```{r, echo=TRUE, eval=FALSE}

tm_shape(gw_ss_15$SDF) +
  tm_fill(col=colnames(gw_ss_15$SDF@data[137]), title="gwr coefficients", style="cont",palette="-RdBu", size=0.2) + 
  tm_layout(
     main.title="GW Correlation to Labour 2015 Vote Share",
     frame=FALSE,
     title.snap.to.legend=FALSE,
     title.size=1,
     title.position = c("left", "top"),
     #inner.margins = c(0,0,0.15,0),
     legend.title.size=1,
     legend.text.size=0.6,
     legend.outside=TRUE)

tm_shape(gw_ss_15$SDF) +
  tm_fill(col=colnames(gw_ss_15$SDF@data[138:147]), title="gwr coefficients", style="cont",palette="-RdBu", size=0.2) + 
  tm_facets(free.scales = FALSE) +
  tm_layout(
     main.title="GW Correlation Between Labour 2015 Vote Share\nand 2017 Swing towards Labour",
     panel.labels=c("Younger Adults", "Degree Educated", "Professionals", "White", "Christian", 
                    "English Speaking", "Single Ethnicity Household", "Social Housing", "No Car", "Not Good Health"),
     frame=FALSE,
     title.snap.to.legend=FALSE,
     title.size=1,
     title.position = c("left", "top"),
     #inner.margins = c(0,0,0.15,0),
     legend.title.size=1,
     legend.text.size=0.6,
     legend.outside=TRUE)

tm_shape(gw_ss_15$SDF) +
  tm_fill(col=colnames(gw_ss_15$SDF@data[128:137]), title="gwr coefficients", style="cont",palette="-RdBu", size=0.2) + 
  tm_facets(free.scales = FALSE) +
  tm_layout(
     main.title="GW Correlation to 2017 Swing to Labour",
     panel.labels=c("Younger Adults", "Degree Educated", "Professionals", "White", "Christian", 
                    "English Speaking", "Single Ethnicity Household", "Social Housing", "No Car", "Not Good Health"),
     frame=FALSE,
     title.snap.to.legend=FALSE,
     title.size=1,
     title.position = c("left", "top"),
     #inner.margins = c(0,0,0.15,0),
     legend.title.size=1,
     legend.text.size=0.6,
     legend.outside=TRUE)
```

## Sinlge GW plot

```{r, echo=TRUE, eval=FALSE}

tm_shape(gw_ss_15$SDF) +
  tm_fill(col=colnames(gw_ss_15$SDF@data[127:147]), title="gwr coefficients", style="cont",palette="-RdBu", size=0.2) + 
  tm_facets(free.scales = FALSE) +
  tm_layout(
     main.title="Geographically Weighted Correlation to 2015 Vote Share and 2017 Swing to Labour",
     panel.labels=c("2015 Support vs\n2017 Swing", "2015 Support vs\nYounger Adults", "2015 Support vs\nDegrees", 
                    "2015 Support vs\nProfessionals", "2015 Support vs\nWhite", "2015 Support vs\nChristian", 
                    "2015 Support vs\nEnglish Speaking", "2015 Support vs\nEthnicity", 
                    "2015 Support vs\nSocial Housing", "2015 Support vs\nNo Car", "2015 Support vs\nNot Good Health", 
                    "2017 Swing vs\nYounger Adults", "2017 Swing vs\nDegrees", "2017 Swing vs\nProfessionals", 
                    "2017 Swing vs\nWhite", "2017 Swing vs\nChristian", "2017 Swing vs\nEnglish Speaking", 
                    "2017 Swing vs\nEthnicity", "2017 Swing vs\nSocial Housing", 
                    "2017 Swing vs\nNo Car", "2017 Swing vs\nNot Good Health"),
     panel.label.size=1.5,
     panel.label.bg.color = "white",
     panel.label.height = 1.25,
     frame=FALSE,
     title.snap.to.legend=FALSE,
     title.size=1,
     title.position = c("left", "top"),
     #inner.margins = c(0,0,0.15,0),
     legend.title.size=1.5,
     legend.text.size=1,
     legend.outside=TRUE)
```

